---
title: "ATE Estimation using Full Propensity Matching and Linear Propensity Scoring"
output: html_notebook
---


```{r}
if(!require("dplyr")){
  install.packages("dplyr")
}

library(dplyr)
```


First, let's load the data and summarize it.
```{r}
lowdim_data <- read.csv("../data/lowDim_dataset.csv")
highdim_data <- read.csv("../data/highDim_dataset.csv")

print(dim(lowdim_data))
print(dim(highdim_data))
```

We can see that the high-dimensional data has an order of magnitude more dimensions, and four times the data as compared to the low dimensional data. Let's view a couple rows

```{r}
head(lowdim_data)
head(highdim_data)
```

Interestingly, the low-dimensional data seems somewhat sparse, with lots of zeros in the feature matrix. Let us proceed with our estimation of the average treatment effect for each.


First, we must train models to estimate the propensity scores:
**TODO: Add all five methods
```{r}
source("../lib/LR_propensity_est.R")

lowdim_model <- lr_propensity_model(lowdim_data)
highdim_model <- lr_propensity_model(highdim_data)
```


Next, we must obtain the propensity score estimates for each dataset:
```{r}
lowdim_prop_scores <- lr_propensity(lowdim_data, lowdim_model)
highdim_prop_scores <- lr_propensity(highdim_data, highdim_model)

```


Now, we must construct a propensity matrix U with all of the information we'll need later; each row will comprise of (index, propensity score, treatment_indicator)
```{r}
lowdim_n <- dim(lowdim_data)[1]
lowdim_index <- c(1:lowdim_n)
lowdim_treatment_indicator <- lowdim_data$A == 1
lowdim_U <- as.matrix(cbind(lowdim_index, lowdim_prop_scores, lowdim_treatment_indicator))

highdim_n <- dim(highdim_data)[1]
highdim_index <- c(1:highdim_n)
highdim_treatment_indicator <- highdim_data$A == 1
highdim_U <- as.matrix(cbind(highdim_index, highdim_prop_scores, highdim_treatment_indicator))
```


In addition, we must construct a matrix with all of the pariwise propensity distances between the points. We will use this for our propensity matching algorithm.

```{r}
source("../lib/distance_matrix.R")

lowdim_D <- get_distance_matrix(lowdim_U)
highdim_D <- get_distance_matrix(highdim_U)
```


Now, we obtain the matching assignments via our Full Matching algorithm (details in Piazza post 118).

```{r}
source("../lib/propensity_fullmatching.R")

lowdim_S <- propensity_matching_root(lowdim_U, lowdim_D)
highdim_S <- propensity_matching_root(highdim_U, highdim_D)
```














