---
title: "ATE Estimation using Full Propensity Matching (implemented from scratch) and Linear Propensity Scoring"
output: html_notebook
---


```{r}
if(!require("dplyr")){
  install.packages("dplyr")
}

library(dplyr)
```


First, let's load the data and summarize it.
```{r}
lowdim_data <- read.csv("../data/lowDim_dataset.csv")
highdim_data <- read.csv("../data/highDim_dataset.csv")

print(dim(lowdim_data))
print(dim(highdim_data))
```

We can see that the high-dimensional data has an order of magnitude more dimensions, and four times the data as compared to the low dimensional data. Let's view a couple rows

```{r}
head(lowdim_data)
head(highdim_data)
```

Interestingly, the low-dimensional data seems somewhat sparse, with lots of zeros in the feature matrix. Let us proceed with our estimation of the average treatment effect for each.

To test our algorithm, let's use a small sample of the datasets
```{r}
lowdim_data <- lowdim_data[1:11,]
highdim_data <- highdim_data[1:11,]
```



First, we must train models to estimate the propensity scores:
**TODO: Add all five methods
```{r}
source("../lib/LR_propensity_est.R")
lowdim_p1_time <- system.time({lowdim_model <- lr_propensity_model(lowdim_data);})
highdim_p1_time <- system.time({highdim_model <- lr_propensity_model(highdim_data);})
```


Next, we must obtain the propensity score estimates for each dataset:
```{r}
lowdim_p1scores_time <- system.time({lowdim_prop_scores <- lr_propensity(lowdim_data, lowdim_model);})
highdim_p1scores_time <- system.time({highdim_prop_scores <- lr_propensity(highdim_data, highdim_model);})

```


Now, we must construct a propensity matrix U with all of the information we'll need later; each row will comprise of (index, propensity score, treatment_indicator)
```{r}
lowdim_n <- dim(lowdim_data)[1]
lowdim_index <- c(1:lowdim_n)
lowdim_treatment_indicator <- lowdim_data$A == 1
lowdim_U <- as.matrix(cbind(lowdim_index, lowdim_prop_scores, lowdim_treatment_indicator))

highdim_n <- dim(highdim_data)[1]
highdim_index <- c(1:highdim_n)
highdim_treatment_indicator <- highdim_data$A == 1
highdim_U <- as.matrix(cbind(highdim_index, highdim_prop_scores, highdim_treatment_indicator))
```


In addition, we must construct a matrix with all of the pariwise propensity distances between the points. We will use this for our propensity matching algorithm.

```{r}
source("../lib/distance_matrix.R")
lowdim_D_time <- system.time({lowdim_D <- get_distance_matrix(lowdim_U);})
highdim_D_time <- system.time({highdim_D <- get_distance_matrix(highdim_U);})
```


Now, we obtain the matching assignments via our Full Matching algorithm (details in Piazza post 118).

```{r}
source("../lib/propensity_fullmatching.R")

lowdim_S_time <- system.time({lowdim_S <- propensity_matching_root(lowdim_U, lowdim_D);})
highdim_S_time <- system.time({highdim_S <- propensity_matching_root(highdim_U, highdim_D);})
```

Finally, we use the assignments to estimate ATE for each dataset, and compare them to the real ATEs

```{r}
lowdim_ATE_real <- 2.0901
highdim_ATE_real <- -54.8558

source("../lib/fullmatch_ATE_est.R")

lowdim_y <- lowdim_data$Y
highdim_y <- highdim_data$Y

lowdim_ATE_est <- estimate_ATE(lowdim_S, lowdim_y)
highdim_ATE_est <- estimate_ATE(highdim_S, highdim_y)
```

```{r}
print("Low Dimensional ATE Estimate: ")
print(lowdim_ATE_est[1])
print("Actual Low Dimensional ATE: ")
print(lowdim_ATE_real[1])
cat("\n")
print("High Dimensional ATE Estimate: ")
print(highdim_ATE_est[1])
print("Actual High Dimensional ATE: ")
print(highdim_ATE_real[1])
```

And below are the times taken for each part of the A1 + D3 + P1 algorithm for ATE estimation

```{r}
cat("Low Dimensional Dataset:")
cat("Time to train LR model=", lowdim_p1_time[3], "s \n")
cat("Time to estimate propensity scores=", lowdim_p1scores_time[3], "s \n")
cat("Time to get pairwise linear propensity distances between all points=", lowdim_D_time[3], "s \n")
cat("Time to obtain full matching over data=", lowdim_S_time[3], "s \n")

cat("\n")

cat("High Dimensional Dataset:")
cat("Time to train LR model=", highdim_p1_time[3], "s \n")
cat("Time to estimate propensity scores=", highdim_p1scores_time[3], "s \n")
cat("Time to get pairwise linear propensity distances between all points=", highdim_D_time[3], "s \n")
cat("Time to obtain full matching over data=", highdim_S_time[3], "s \n")
```













